---
title: "Practical Machine Learning - Prediction Assignment Writeup"
author: "Michael Lange"
date: "22. November 2015"
output: html_document
---

### Background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement and a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).


### Data
The training data for this project are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har.


### Reproducibility
```{r}
library(caret)
library(randomForest)
library(rpart)
library(rattle)

set.seed(4711)
```


### Loading and Cleaning Data
In this section, after loading the data, columns with missing values and irrelevant variables will be delected.

```{r}
trainUrl <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testUrl <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

training <- read.csv(url(trainUrl), na.strings=c("NA","#DIV/0!",""))
testing <- read.csv(url(testUrl), na.strings=c("NA","#DIV/0!",""))

training1 <- training[,colSums(is.na(training)) == 0]
testing1 <- testing[,colSums(is.na(testing)) == 0]

# Deleting irrelevant variables (user_name, raw_timestamp_part_1, raw_timestamp_part_, 2 cvtd_timestamp, new_window, and num_window; columns 1 to 7)
training1 <- training1[,-c(1:7)]
testing1 <- testing1[,-c(1:7)]

dim(training1)
dim(testing1)
```


### Partioning the training set into two
```{r}
inTrain <- createDataPartition(y = training1$classe, p = 0.75, list = FALSE)

myTraining <- training1[inTrain,]
myTesting <- training1[-inTrain,]

dim(myTraining)
dim(myTesting)
```


### Prediction Model - Random Forest
```{r}
modFit <- randomForest(classe ~ ., data = myTraining, method = "class")

prediction <- predict(modFit, myTesting, type = "class")

confusionMatrix(prediction, myTesting$classe)
```
As seen by the result of the confusionmatrix, the model is good and efficient because it has an accuracy of 0.9947 and very good sensitivity & specificity values on the testing dataset. (the lowest value is 0.9888 for the sensitivity of the class D)


### Creating submission files
```{r}
predictionfinal <- predict(modFit, testing, type = "class")
predictionfinal

pml_write_files = function(x){
    n = length(x)
    for(i in 1:n){
        filename = paste0("problem_id_",i,".txt")
        write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
    }
}

#pml_write_files(prediction)
```

